### 2025-04-19 manuscript/dissertation.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  abstract=true]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=30mm,left=30mm,right=30mm,heightrounded]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% TODO: Add custom LaTeX header directives here
\usepackage{float, booktabs, siunitx, orcidlink}
\setkomafont{disposition}{\bfseries}
\providecommand{\keywords}[1]
{
  \small	
  \textbf
{\textit{Keywords---}} #1
}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={European or American? A re-examination of the transatlantic influence over the IMF},
  pdfkeywords={IMF, US hegemony, Europe},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{European or American? A re-examination of the transatlantic
influence over the IMF}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{GV499 Dissertation}
\author{Dianyi Yang~\orcidlink{0009-0004-4652-3429}\textsuperscript{}}
\date{2024-01-06}

\begin{document}
\maketitle
\begin{abstract}
This is an abstract.
\end{abstract}
\begin{center}
\keywords{IMF, US hegemony, Europe}
\end{center}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The International Monetary Fund (IMF) and the World Bank\footnote{Familar
  name for the International Bank for Reconstruction and Development
  (IBRD), the lending arm of the later created World Bank Group.} were
established toward the end of World War II as part of the Bretton Woods
agreement in 1944 (\protect\hyperlink{ref-jensen2004}{Jensen, 2004}).
While the World Bank specialised in long-term project financing, the IMF
was initially designed for addressing the short-term balance-of-payment
crises which would otherwise undermine the fixed exchange rate
arrangements under the Bretton Woods system
(\protect\hyperlink{ref-jensen2004}{Jensen, 2004}).

One renowned fact about the IMF and its sister institution, the World
Bank, is the long-standing ``gentlemen's agreement'' of the IMF managing
director being a European, and the World Bank president being an
American (\protect\hyperlink{ref-Keating2024}{Keating, 2024};
\protect\hyperlink{ref-Weiss2023}{Weiss, 2023}), despite no such
provisions in either institution's articles of agreements
(\protect\hyperlink{ref-Fund2020}{IMF, 2020};
\protect\hyperlink{ref-WorldBank2021}{World Bank, 2021}).

Such arrangements inspire a straightforward question that is yet
difficult to answer: who has more influence over the IMF, the US or
Europe? Whereas the US dominance in the World Bank is undisputed
(\protect\hyperlink{ref-clark2020}{Clark \& Dolan, 2020};
\protect\hyperlink{ref-wade2002}{Wade, 2002};
\protect\hyperlink{ref-woods2003}{Woods, 2003}), the case of the IMF is
much more complicated. As the
\protect\hyperlink{sec-litreview}{Literature Review} section
demonstrates, there has been abundant evidence of both European and
American influence on the IMF, but short of a qualitative or
quantitative comparison between the two.

Qualitative attempts at this question may have been thwarted by the
closed-door and consensus-based style of governance of the IMF, as
epitomised by the ``gentlemen's agreement''. On the one hand, the
frequent use of conventions and consensus renders any analyses on the
formal structure of the IMF (e.g.~veto power) insufficient on its own.
On the other hand, the closed-door negotiations imply that we often
observe the outcome, but not the process of the bargaining. While
specific case studies (e.g.~Mexico and European debt crisis, see Lipscy
\& Lee (\protect\hyperlink{ref-lipscy2018}{2018}) and Hodson
(\protect\hyperlink{ref-hodson2014}{2014})) may have shed light on the
influence of the US or Europe, they rarely allow for a transatlantic
comparison and their representativeness can be questioned.

This dissertation therefore attempts to quantitatively compare the
influence of the US and Europe on the IMF.

The structure of this dissertation is as follows:
Section~\ref{sec-litreview} provides a review over the literature on the
influence of the US and Europe over the IMF.

\hypertarget{sec-litreview}{%
\section{Literature Review}\label{sec-litreview}}

\begin{itemize}
\tightlist
\item
  IMF chief European (\protect\hyperlink{ref-Kahler2001}{Kahler, 2001})
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-clark2020}{}}%
Clark, R., \& Dolan, L. R. (2020). Pleasing the Principal: U.S.
Influence in World Bank Policymaking. \emph{American Journal of
Political Science}, \emph{65}(1), 36--51.
\url{https://doi.org/10.1111/ajps.12531}

\leavevmode\vadjust pre{\hypertarget{ref-hodson2014}{}}%
Hodson, D. (2014). The IMF as a de facto institution of the EU: A
multiple supervisor approach. \emph{Review of International Political
Economy}, \emph{22}(3), 570--598.
\url{https://doi.org/10.1080/09692290.2014.956136}

\leavevmode\vadjust pre{\hypertarget{ref-Fund2020}{}}%
IMF. (2020). \emph{{Articles} of {Agreement}}. International Monetary
Fund.

\leavevmode\vadjust pre{\hypertarget{ref-jensen2004}{}}%
Jensen, N. M. (2004). Crisis, Conditions, and Capital. \emph{Journal of
Conflict Resolution}, \emph{48}(2), 194--210.
\url{https://doi.org/10.1177/0022002703262860}

\leavevmode\vadjust pre{\hypertarget{ref-Kahler2001}{}}%
Kahler, M. (2001). \emph{Leadership selection in the major
multilaterals}. Institute for International Economics.

\leavevmode\vadjust pre{\hypertarget{ref-Keating2024}{}}%
Keating, J. E. (2024). Why {Is} the {IMF} {Chief} {Always} a {European}?
In \emph{Foreign Policy}.
\url{https://foreignpolicy.com/2011/05/18/why-is-the-imf-chief-always-a-european/}

\leavevmode\vadjust pre{\hypertarget{ref-lipscy2018}{}}%
Lipscy, P. Y., \& Lee, H. N.-K. (2018). The IMF As a Biased Global
Insurance Mechanism: Asymmetrical Moral Hazard, Reserve Accumulation,
and Financial Crises. \emph{International Organization}, \emph{73}(1),
35--64. \url{https://doi.org/10.1017/s0020818318000371}

\leavevmode\vadjust pre{\hypertarget{ref-wade2002}{}}%
Wade, R. H. (2002). US hegemony and the World Bank: the fight over
people and ideas. \emph{Review of International Political Economy},
\emph{9}(2), 215--243. \url{https://doi.org/10.1080/09692290110126092}

\leavevmode\vadjust pre{\hypertarget{ref-Weiss2023}{}}%
Weiss, M. A. (2023). Selecting the world bank president. In
\emph{Congressional Research Service}.
\url{https://sgp.fas.org/crs/row/R42463.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-woods2003}{}}%
Woods, N. (2003). \emph{The united states and the international
financial institutions: Power and influence within the world bank and
the IMF} (pp. 92--114). Oxford University PressOxford.
\url{https://doi.org/10.1093/0199261431.003.0005}

\leavevmode\vadjust pre{\hypertarget{ref-WorldBank2021}{}}%
World Bank. (2021). \emph{Articles of agreement}.
\url{https://www.worldbank.org/en/about/articles-of-agreement}

\end{CSLReferences}

\end{document}