### 2025-04-22 demo/demo.qmd
fix demo paths (#3)

---
title: European or American? A re-examination of the transatlantic influence over the IMF
# subtitle: GV499 Dissertation
header-includes: |
  \AtBeginDocument{\title[GV499 Dissertation]{European or American? A re-examination of the transatlantic influence over the IMF}}
date: last-modified
author: Dianyi Yang
format:
  beamer:
    theme: Madrid
bibliography: references.bib
csl: extra/apa.csl
project:
  execute-dir: "D:/OneDrive - London School of Economics/Desktop/GV499/IMF"
---

## Background

-   The International Monetary Fund (IMF) and the World Bank were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004].

-   Believed US dominance in the beginning,

-   "Gentlemen's agreement": the IMF managing director being a European, and the World Bank president being an American [@Keating2024; @Weiss2023].

## Research Question

Who has more influence over the IMF, the US or Europe?

-   US?

    -   HQ in the US

    -   US veto power

    -   US hegemony

    -   US deputy

    -   Case Study: Mexico

-   Europe?

    -   European managing director

    -   Case Study: Euro Crisis

-   Difficulties in comparison

    -   Consensus-based governance

    -   closed-door negotiations

## Empirical Strategy

-   Data source/empirical inspiration: @lipscy2018 (1980-2010)

    -   but their focus is the effect on self-insurance

    -   did not separate Europe from the US - 'Western influence'

-   Proxies for influence (merged through PCA)

    -   IMF direct: nationals employed as economists, IMF quota.

    -   US: UN affinity score with US, trade volume with US, US bank lending

    -   European: UN affinity score with European powers, trade volume with European powers, European bank lending

## Empirical Strategy

-   Dependent variables

    -   IMF loan to GDP ratio

    -   IMF participation rate

    -   IMF loan approval

    -   IMF conditions

-   Independent variables

    -   IMF influence (used by @lipscy2018 and separated by me into EU, US and IMF)

    -   GDP

    -   Growth

    -   Reserves

    -   OECD

-   Hypotheses

    -   $H_0$: No difference in the influence of US and Europe over the IMF.

    -   $H_1$: Either the US or EU has more influence over the IMF.

    -   Test of equality of the US and EU coefficients.

## Results from [@lipscy2018]

![](images/clipboard-2567109775.png)

## Preliminary Results

```{r setup}
need <- c('tidyverse','haven','AER','fixest','modelsummary',
          'corrplot', 'kableExtra','car','parameters') # list packages needed
have <- need %in% rownames(installed.packages()) # checks packages you have
if(any(!have)) install.packages(need[!have]) # install missing packages
invisible(lapply(need, library, character.only=T))
```

```{r results}
load('../save/regtable.RData')

#t-tests
regtable$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(regtable$`Probit: IMF loan approval`)

glance_custom.tobit <- function(x, ...) {
  data.frame(
    'equality'=paste0('[','$p=',round(lht(x, test='F', 'us=eu')[2,4],3),'$]'),
    'vcov.type' = 'by: Country'
  )
}

glance_custom.fixest <- function(x, ...) {
  data.frame(
    'equality'=paste0('[','$p=',round(lht(x, test='F', 'us=eu')[2,4],3),'$]'),
    'vcov.type' = 'by: Country'
  )
}

coefmap <- c('us' = 'US Influence',
             'eu' = 'EU Influence',
             'imf' = 'IMF Influence',
             'lnrgdpnew' = 'GDP',
             'lnrgdpnewsq' = 'GDP$^2$',
             'rgdpchnew' = 'GDPpc',
             'rgdpchnewsquare' = "GDPpc$^2$",
             'growth1new' = 'GDPpc growth',
             'reserv1' = 'Reserves',
             'oecd1' = 'OECD',
              '(Intercept)' = '(Intercept)')
gof_map <- list(list("raw" = "equality", "clean" = "US=EU", "fmt" = NULL),
                list("raw" = "nobs", "clean" = "$N$", "fmt" = 0),
                list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL))
msummary(regtable, coef_map=coefmap, gof_map=gof_map, stars=T, escape=F, output = "kableExtra") %>% 
  kable_styling(latex_options="scale_down")
```

## References {.unnumbered}

::: {#refs}
:::