### 2025-04-22 demo/demo.qmd
Make table in PCA.R (with tinytable)

---
title: European or American? A re-examination of the transatlantic influence over the IMF
header-includes: |
  \AtBeginDocument{\title[European or American]{European or American? A re-examination of the transatlantic influence over the IMF}}
date: last-modified
author:
- name: Dianyi Yang
# - name: Tong Hu // name placeholders subject to change
# - name: Zhenyi Chen
format:
  beamer:
    theme: Madrid
    include-in-header: 
      - text: |
          \definecolor{myblue}{RGB}{0, 113, 187}
          \colorlet{beamer@blendedblue}{myblue}
---

## Background

-   The International Monetary Fund (IMF) and the World Bank were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004].

-   Believed US dominance in the beginning,

-   "Gentlemen's agreement": the IMF managing director being a European, and the World Bank president being an American [@Keating2024; @Weiss2023].

## Research Question

Who has more influence over the IMF, the US or Europe?

-   US?

    -   HQ in the US

    -   US veto power

    -   US hegemony

    -   US deputy

    -   Case Study: Mexico

-   Europe?

    -   European managing director

    -   Case Study: Euro Crisis

-   Difficulties in comparison

    -   Consensus-based governance

    -   closed-door negotiations

## Empirical Strategy

-   Data source/empirical inspiration: @lipscy2018 (1980-2010)

    -   but their focus is the effect on self-insurance

    -   did not separate Europe from the US - 'Western influence'

-   Proxies for influence (merged through PCA)

    -   IMF direct: nationals employed as economists, IMF quota.

    -   US: UN affinity score with US, trade volume with US, US bank lending

    -   European: UN affinity score with European powers, trade volume with European powers, European bank lending

## Empirical Strategy

-   Dependent variables

    -   IMF loan to GDP ratio, IMF participation rate

    -   IMF loan approval, IMF conditions

-   Independent variables

    -   IMF influence (used by @lipscy2018 and separated by me into EU, US and IMF)

    -   GDP, Growth

    -   Reserves, OECD

-   Hypotheses

    -   $H_0$: No difference in the influence of US and Europe over the IMF.

    -   $H_1$: Either the US or EU has more influence over the IMF.

    -   Test of equality of the US and EU coefficients.

## Results from [@lipscy2018]

![](images/clipboard-2567109775.png)

## Preliminary Results

```{r setup}
need <- c('AER', 'car', 'kableExtra', 'modelsummary', 'parameters') # list packages needed
# have <- need %in% rownames(installed.packages()) # checks packages you have
# if(any(!have)) install.packages(need[!have]) # install missing packages
invisible(lapply(need, library, character.only = T))
```

```{r results}
load('save/regTable.RData')
resultsTable
```

## References {.unnumbered}

::: {#refs}
:::