### 2025-05-18 manuscript/dissertation.qmd
correct manuscript "familiar" typo

---
title: European or American? A re-examination of the transatlantic influence over the IMF
# subtitle: GV499 Dissertation
date: last-modified
authors:
  - name: Dianyi Yang
    email: d.yang17@lse.ac.uk
    orcid: 0009-0004-4652-3429
    correspondence: true
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
abstract: This is an abstract.
keywords: IMF, US hegemony, Europe
---

\newpage

# Introduction

The International Monetary Fund (IMF) and the World Bank[^1] were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004]. While the World Bank specialised in long-term project financing, the IMF was initially designed for addressing the short-term balance-of-payment crises which would otherwise undermine the fixed exchange rate arrangements under the Bretton Woods system [@jensen2004].

[^1]: Familiar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.

One renowned fact about the IMF and its sister institution, the World Bank, is the long-standing "gentlemen's agreement" of the IMF managing director being a European, and the World Bank president being an American [@Keating2024; @Weiss2023], despite no such provisions in either institution's articles of agreements [@Fund2020; @WorldBank2021].

Such arrangements inspire a straightforward question that is yet difficult to answer: who has more influence over the IMF, the US or Europe? Whereas the US dominance in the World Bank is undisputed [@clark2020; @wade2002; @woods2003], the case of the IMF is much more complicated. As the [Literature Review](#sec-litreview) section demonstrates, there has been abundant evidence of both European and American influence on the IMF, but short of a qualitative or quantitative comparison between the two.

Qualitative attempts at this question may have been thwarted by the closed-door and consensus-based style of governance of the IMF, as epitomised by the "gentlemen's agreement". On the one hand, the frequent use of conventions and consensus renders any analyses on the formal structure of the IMF (e.g. veto power) insufficient on its own. On the other hand, the closed-door negotiations imply that we often observe the outcome, but not the process of the bargaining. While specific case studies (e.g. Mexico and European debt crisis, see @lipscy2018 and @hodson2014) may have shed light on the influence of the US or Europe, they rarely allow for a transatlantic comparison and their representativeness can be questioned.

This dissertation therefore attempts to quantitatively compare the influence of the US and Europe on the IMF.

The structure of this dissertation is as follows: @sec-litreview provides a review over the literature on the influence of the US and Europe over the IMF.

# Literature Review {#sec-litreview}

-   IMF chief European [@Kahler2001]

# References {.unnumbered}

::: {#refs}
:::