### 2025-05-27 demo/demo.qmd
Add interpretation

---
title: European or American? A re-examination of the transatlantic influence over the IMF
header-includes: |
  \AtBeginDocument{\title[European or American]{European or American? A re-examination of the transatlantic influence over the IMF}}
date: last-modified
author:
- name: Dianyi Yang
# - name: Tong Hu // name placeholders subject to change
# - name: Zhenyi Chen
format:
  beamer:
    theme: Madrid
    link-citations: true
    include-in-header: 
      - text: |
          \definecolor{myblue}{RGB}{0, 113, 187}
          \colorlet{beamer@blendedblue}{myblue}
---

## Background

-   The International Monetary Fund (IMF) and the World Bank were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004].

-   Believed US dominance in the beginning,

-   "Gentlemen's agreement": the IMF managing director being a European, and the World Bank president being an American [@keating2024; @weiss2023].

## Research Question

Who has more influence over the IMF, the US or Europe?

-   US?

    -   HQ in the US

    -   US veto power

    -   US hegemony

    -   US deputy

    -   Case Study: Mexico

-   Europe?

    -   European managing director

    -   Case Study: Euro Crisis

-   Difficulties in comparison

    -   Consensus-based governance

    -   closed-door negotiations

## Empirical Strategy

-   Data source/empirical inspiration: @lipscy2018 (1980-2010)

    -   but their focus is the effect on self-insurance

    -   did not separate Europe from the US - 'Western influence'

-   Proxies for influence (merged through PCA)

    -   IMF direct: nationals employed as economists, IMF quota.

    -   US: UN affinity score with US, trade volume with US, US bank lending

    -   European: UN affinity score with European powers, trade volume with European powers, European bank lending

## Empirical Strategy

-   Dependent variables

    -   IMF loan to GDP ratio, IMF participation rate

    -   IMF loan approval, IMF conditions

-   Independent variables

    -   IMF influence (used by @lipscy2018 and separated by me into EU, US and IMF)

    -   GDP, Growth

    -   Reserves, OECD

-   Hypotheses

    -   $H_1$: Either the US or Europe has more influence over the IMF.

        -   Test of equality of the US and EU coefficients.

    -   $H_2$: The US/Europe exerts a statistically significant influence on IMF lending decisions, controlling for European/American influence.

## Results from [@lipscy2018]

![](images/clipboard-2567109775.png)

## Preliminary Results

```{r results}
load('save/regTable.RData')
resultsTable |> tinytable::theme_tt("resize", width = 0.81)
```

## Interpretation of Results

-   $H_1$: The evidence supports the hypothesis that Europe exerts greater influence over the IMF than the United States, particularly in determining loan amounts and participation rates.

-   $H_2$: The results offer a more nuanced view.

    -   European influence is statistically significant in shaping the loan-to-GDP ratio, participation rate, and loan approval.

    -   In contrast, US influence is **only** significant in reducing the number of IMF conditions.

    -   This finding echoes @lutz2014, who argue that European actors have historically adhered more strictly to the principles of the Washington Consensus than their US counterparts.

## References {.unnumbered}

::: {#refs}
:::