### 2025-07-29 manuscript/manuscript.qmd
EU lit review revised_0729

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
authors:
  - name: Dianyi Yang
    email: d.yang17@lse.ac.uk
    orcid: 0009-0004-4652-3429
    correspondence: true
  - name: Tong Hu
    orcid: 0009-0003-5234-8164
  - name: Zhenyi Chen
    orcid: 0009-0008-2910-5187
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
abstract: This is an abstract.
keywords: IMF, US hegemony, Europe
execute:
  echo: false
  warning: false
  error: false
  message: false
---

\newpage

# Introduction

The International Monetary Fund (IMF) and the World Bank[^1] were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004]. While the World Bank specialised in long-term project financing, the IMF was initially established for providing short-term balance-of-payment support and ensuring the stability of the international monetary system  [@fund2020; @jensen2004].

[^1]: Familiar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.

Despite IMF's initial technical and narrow mandate, the IMF gradually expanded its role into structural adjustment and domestic economic governance, with loan conditionality often targeting fiscal policy, labour markets, and institutional reforms [@barnett2004]. This evolution pushed the Fund into deeply political and distributional arenas far beyond its original design.

However, despite the far-reaching nature of IMF loan conditionality—often inspired by the tenets of neoclassical economics [@best2007; @clift2021; @chwieroth2015; @nelson2014], such as fiscal austerity, market liberalisation, and privatisation—there is little robust evidence that these prescriptions consistently lead to improved economic outcomes. In many cases, IMF programmes have been associated with worsening economic inequality [@garuda2000; @lang2020; @oberdabernig2013], stagnating growth [@dreher2006; @przeworski2000; @oberdabernig2013], or social unrest [@hartzell2010; @casper2017], raising doubts about the effectiveness and legitimacy of its policy advice.

This performance problem becomes a *legitimacy* issue when compounded by the Fund’s institutional design, which structurally privileges the interests of its most powerful member states [@stone2011]. With a weighted voting system [@fund2020] and governance norms that concentrate decision-making among a handful of advanced economies—particularly those in the West—the IMF has long been criticised as a platform for great power projection [@barnett2004]. Scholars have drawn parallels between the Fund’s influence and forms of informal empire or *neo-colonialism*, where powerful donor states impose conditions on economically weaker recipients through seemingly technocratic institutions [@cain2022; @chen2021; @hickel2020; @moyo2024].

While critiques of Western dominance in the IMF are longstanding, they often treat “the West” as a monolithic bloc. In practice, however, the institutional structure of the IMF reflects a careful transatlantic balancing act. One of the most symbolic manifestations of this arrangement is the long-standing “gentlemen’s agreement” under which the IMF’s Managing Director is always a European, while the World Bank is headed by an American [@keating2024; @weiss2023]. This tacit division of leadership—despite being absent from the formal Articles of Agreement [@fund2020; @WorldBank2021]—raises an underexplored but important question: *who actually wields more influence within the IMF, Europe or the United States?* Addressing this question helps clarify the nature of global economic governance and sheds light on the dynamics within what is often treated as a unified Western core.

The question is straightforward but surprisingly difficult to answer. Whereas the US dominance in the World Bank is undisputed [@clark2020; @wade2002; @woods2003], the IMF presents a more ambiguous case. As the [Literature Review](#sec-litreview) section shows, existing research offers abundant evidence of both American and European influence, yet rarely attempts a direct comparison between the two.

Qualitative approaches to this issue are often hampered by the Fund’s closed-door, consensus-based decision-making [@stone2011, pp.55-56]—precisely the kind of opaque process exemplified by the “gentlemen’s agreement.” On the one hand, the prevalence of informal conventions limits the utility of formal institutional analyses (e.g., voting power or veto rights). On the other, the secretive nature of negotiations means scholars often observe the outcomes, but not the underlying power struggles. While case studies such as Mexico’s bailout or the European debt crisis [@lipscy2018; @copelovitch2010a; @hodson2014] provide compelling narratives, they lack transatlantic comparability and generalisability.

This paper contributes to the literature in three key ways. First, it offers a direct empirical comparison of US and European influence over IMF lending outcomes—an angle surprisingly overlooked in existing work. Second, by disaggregating “the West” into its US and European components, it uncovers variation in how influence is exercised and to what ends. Third, it challenges the assumption of geographic bias, showing that European powers do not disproportionately favour European borrowers. These findings shed new light on the internal politics of the IMF and speak to wider debates on institutional legitimacy, great power contestation, and the geopolitics of global finance.

The structure of this paper is as follows: @sec-litreview provides a review over the literature on the influence of the US and Europe over the IMF.

# Literature Review {#sec-litreview}

International Relations theory, particularly realism, has long emphasised that international organisations reflect underlying distributions of power rather than operating as neutral or autonomous actors. As Mearsheimer famously argued, international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994]. The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a; @stone2011; @vreeland2009]. Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.

In the case of the IMF, there is broad consensus that the institution is dominated by “Western” powers. This is reflected in both historical accounts of its founding and institutional design—especially its weighted voting system, which concentrates power in the hands of the United States and European countries [@kahler2001; @barnett2004; @copelovitch2010]. Given the ideological, geopolitical, and economic proximity between the US and major European states, many studies treat “the West” as a unified bloc and do not disaggregate their respective influences [e.g. @copelovitch2010a; @lipscy2018; @barro2005].

For the strand of literature that examines American and European influence separately, nearly all studies identify the United States as the dominant actor within the IMF [@stone2011, p.51]. Qualitative accounts emphasise the United States’ foundational role in shaping the IMF’s governance structure [@stone2011], the Fund’s practice of making direct enquiries with Washington for major lending decisions [@southard1979; @woods2003], and its unique status as the only country-level veto player, stemming from its position as the largest financial contributor [@woods2003; @stone2011]. The US also benefits from significant informational advantages over other members [@stone2011, p.57] and proximity-based access, with the IMF headquartered in Washington, D.C. [@stone2011, p.57]. Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018]. This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].

Compared to the extensive literature on the United States, scholarly analyses of European influence within the IMF remain relatively limited and fragmented. A number of scholars (such as @smaghi2004; @thacker1999; @miescu2014) emphasize that European representation within the Fund is institutionally dispersed, resulting in a fragmented voice that limits collective influence. In contrast, the United States retains a centralized and coherent presence, thereby maintaining a dominant position in key decision-making processes. Nevertheless, other contributions (such as @keating2024; @lutz2014; @copelovitch2010a; @stone2004) argue that European states possess significant structural and institutional advantages—including leadership appointments, voting weight, agenda-setting capacity, transnational financial exposure, and diplomatic leverage—which enable them to exert substantive influence over IMF lending practices.

First, European countries benefit from long-standing institutional privileges within the IMF’s governance architecture. As noted in the introduction, the Fund’s leadership has traditionally been allocated to Europe under an informal transatlantic arrangement that dates back to the Bretton Woods settlement [@kahler2001]. Collectively, EU member states control over 30% of the IMF’s voting rights, with the United Kingdom, France, and Germany each occupying individual executive board seats [@leech2005a]. Quantitative studies suggest that, under a scenario of unified representation, the EU could potentially attain voting power equivalent to a de facto veto similar to that of the United States [@brandner2009a]. These structural features provide European countries with disproportionate influence relative to their global economic weight.

Second, European countries have played a leading role in shaping policy design and operational implementation within the IMF. @lutz2014 demonstrates that the EU and the European Central Bank were instrumental in embedding austerity-oriented conditionalities—reminiscent of the Washington Consensus—in IMF lending to Central and Eastern European states such as Latvia and Romania. This policy orientation was motivated by the significant exposure of French and German banks to these economies. At the request of EU institutions, IMF conditionality in these cases expanded beyond standard macroeconomic stabilization to encompass fiscal rules, structural adjustment, and financial sector supervision [@hodson2014]. Coordination platforms such as EURIMF (European Representatives to the IMF) and SCIMF (Sub-Committee on the IMF) have enabled EU member states to form cohesive negotiating positions and act collectively to advance shared preferences within the Fund [@lutz2014].

In this context, @copelovitch2010 argues that IMF lending decisions are fundamentally political, with European influence manifesting not only through formal mechanisms such as quota shares, but also through informal channels including diplomatic lobbying, technical staffing appointments, and the selection of mission team leaders. During the Eurozone sovereign debt crisis, major European powers, particularly France and Germany, actively solicited IMF participation in stabilization packages. The resulting “troika” mechanism—comprising the IMF, the European Commission, and the European Central Bank—was emblematic of Europe's dual status as both lender and borrower in the crisis [@veron2016]. However, this arrangement attracted criticism for compromising IMF independence. Some scholars contend that excessive deference to European preferences undermined the Fund’s credibility, and that the reluctance of euro area governments to implement IMF-prescribed reforms was a key driver of the crisis’s deepening [@veron2016].

Third, European influence over IMF lending has often served to advance strategic financial and geopolitical interests. This was particularly evident during the Eurozone crisis, as mentioned earlier, when major European economies—most notably Germany and France—sought to mitigate systemic risks arising from their banking sector’s exposure to highly indebted member states. Fiscal surveillance mechanisms, such as the EU’s 3% deficit ceiling, were complemented by IMF-backed assistance programs aimed at restoring financial stability [@heimberger2025, p. 9]. The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request [@kickert2019a]. Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments [@kickert2019a].

Moreover, European states have extended their influence to extra-European contexts, particularly in Africa. Countries such as the United Kingdom and France have supported IMF concessional lending to former colonies, often motivated by a combination of commercial and strategic interests, including access to natural resources and the preservation of post-colonial diplomatic ties [@stone2004]. However, this form of influence has not been without controversy. Critics such as @moyo2024 and @cain2022 argue that European-backed IMF programs have promoted neoliberal policy frameworks—including market liberalization and privatization—without adequate consideration of local development priorities, undermining African countries' economic sovereighty and state capacity.

As noted earlier, numerous scholars have questioned both the extent and effectiveness of Europe’s influence within the IMF. First, Europe faces persistent challenges in internal coordination. EU member states are represented through multiple, uncoordinated constituencies within the Fund, resulting in a fragmented and often inconsistent policy voice [@smaghi2004]. This institutional fragmentation is compounded by substantive policy divergences among key member states—for instance, France tends to prioritize pro-growth strategies, while Germany places greater emphasis on fiscal discipline [@miescu2014].

Second, despite Europe’s substantial voting power, the United States continues to enjoy structural advantages, most notably its effective veto power over major IMF decisions and its ability to shape key institutional outcomes [@thacker1999]. Moreover, European influence is often constrained by the need to coordinate positions with other members of the G5, which may necessitate concessions that dilute European preferences or undermine coherence [@copelovitch2010a].

Third, concerns have been raised regarding the democratic legitimacy and global representativeness of European countries within the IMF governance structure. @woodward2024 highlights the disproportionate relationship between the population size of EU member states and their aggregate voting power in the Fund, arguing that this imbalance structurally disadvantages developing countries. @leech2005a further suggests that while a consolidation of European seats could potentially establish a bipolar governance configuration capable of counterbalancing U.S. dominance, such a shift might also open space for increased participation and influence by small and medium-sized states within the IMF’s decision-making processes.

However, while these studies offer important insights, they tend to remain case-specific or anecdotal. Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.

In principle, quantitative approaches are well suited for directly comparing the influence of different great powers within international organisations. In the IMF context, however, most such studies focus on American influence, with little systematic attention paid to European power. These studies typically proxy the strategic interests of major powers using indicators such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004]. Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009]. Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005]. These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009]. Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.

This paper addresses this gap by conducting a disaggregated, quantitative comparison of US and European influence on IMF lending outcomes. By distinguishing between the two major transatlantic actors, we provide new insights into the internal dynamics of power within the IMF and contribute to broader debates over institutional legitimacy, hegemony, and the geopolitics of economic governance.

# Empirical Analysis

## Hypotheses

Since the [literature review](#sec-litreview) has outlined arguments for both US and European influence over IMF decision-making, this section turns to a quantitative comparison of their respective impacts. Accordingly, we formulate our main hypothesis:

\begin{restatable}[$H_{1}$]{hyp}{hypmain}
\label{hyp-main}
The United States and Europe exert different levels of influence on IMF lending decisions.
\end{restatable}

Moreover, prior literature has either combined the United States and Europe under the umbrella of “Western” influence on IMF lending decisions [@lipscy2018], or has examined the influence of one without accounting for the other [@oatley2004; @thacker1999]. To address this gap, we reassess the influence of the United States and Europe separately, while controlling for the presence of the other. Accordingly, we formulate the following two hypotheses:

```{=latex}
\setcounter{hyp}{2}
\begin{restatable}[$H_{2a}$]{subhyp}{hypus}
\label{hyp-us}
The United States exerts a statistically significant influence on IMF lending decisions, controlling for European influence.
\end{restatable}
\vspace{-1em} % remove vertical space
\begin{restatable}[$H_{2b}$]{subhyp}{hypeu}
\label{hyp-eu}
Europe exerts a statistically significant influence on IMF lending decisions, controlling for United States influence.
\end{restatable}
```

## Data

```{r}
#| label: fig-data
dta <- readr::read_rds('data/panel_data.rds')
```

This paper uses the dataset compiled by @lipscy2018, which extends the @barro2005 dataset through 2010. It provides panel data for 130 countries in five-year intervals from 1980 to 2010[^2], including information on IMF borrowing and participation outcomes, economic indicators, and measures of each country's relationship with the United States and major European powers.

[^2]: e.g. 1980-1985, 1985-1990.

## Dependent Variables

In line with @lipscy2018 and @barro2005, we use the following four measures of IMF lending outcomes as our dependent variables:

-   IMF loan to GDP ratio
-   IMF participation rate
-   IMF loan to GDP approval
-   Number of IMF conditions

These variables follow the definitions used in @lipscy2018. The *loan-to-GDP ratio* captures the average size of IMF lending relative to a country's GDP over each five-year interval. The *participation rate* measures the share of months within each interval during which a country was subject to an IMF programme. The *approval indicator* is a binary variable coded 1 if a new IMF programme was initiated during the interval, and 0 otherwise [see also @barro2005]. Finally, the *number of conditions* reflects the total number of performance criteria, benchmarks, and prior actions required for disbursement under IMF programmes [see also @copelovitch2010; @dreher2015].

## Independent Variables

To assess the influence of the United States and European powers on IMF lending decisions, we construct two key independent variables that capture each actor's *propensity* to exert influence in favour of a country's borrowing from the IMF at a given time. These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018]. This approach assumes that both the US and European powers[^3] are more likely to advocate for lending decisions that benefit their close partners.

[^3]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values. This approach reflects the assumption that smaller European countries are unlikely to exert meaningful influence over IMF decisions on behalf of their partners.

Specifically, we construct our independent variables using three indicators commonly interpreted as channels of political and economic affinity, similar to @lipscy2018:

-   UN General Assembly voting similarity[^4]
-   Bilateral trade intensity[^5]
-   Banking sector exposure[^6]

[^4]: Originallly used in @thacker1999.

[^5]: From @barro2005.

[^6]: See @broz2006; @lipscy2017; @oatley2004.

```{r percentage_variance}
load('save/PCA.Rdata')
US_PC1_percentage <- US$sdev[1]^2 / sum(US$sdev^2) * 100
EU_PC1_percentage <- EU$sdev[1]^2 / sum(EU$sdev^2) * 100
```

However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^7]. To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008]. Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^8].

[^7]: See Appendix [-@sec-corr].

[^8]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States). See Appendix [-@sec-scree] for scree plots and additional details.

Our approach differs from @lipscy2018 in that we do not pool the US and European variables together. Instead, we construct two separate measures: US influence and European influence. The PCA is conducted separately for each actor, using the same three indicators described above. The resulting first principal components are then normalised to have a mean of zero and a standard deviation of one, allowing for easier interpretation and comparability of coefficients in subsequent analyses.

## Covariates

The existing literature has identified several factors beyond geopolitical alignment that influence IMF lending decisions or a country's demand for IMF assistance. We include these variables as covariates to account for potential confounding effects. Notably, countries with larger IMF quotas (and thus greater voting power), as well as those with nationals among IMF staff, are expected to receive more favourable treatment from the Fund [@barro2005].

We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997]. To capture these dynamics, we include GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance. International reserves as a share of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support. Finally, a binary variable for OECD membership serves as a proxy for broader institutional and economic differences among countries.

The selection of covariates, including their polynomial transformations[^9], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.

[^9]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.

## Model Specification {#sec-model}

One notable feature of IMF loan outcomes is that they are *left-censored* at zero—that is, loan amounts, participation rates, and the number of conditions are often zero, but cannot take negative values. In such cases, ordinary least squares (OLS) estimation yields biased and inconsistent results. The Tobit model addresses this issue by explicitly modelling the latent continuous variable ($Y_{it}^*$) underlying the censored observed outcomes ($Y_{it}$) [@tobin1958], and is particularly well-suited to our setting [@amemiya1984].

For each non-binary dependent variable $Y_{it}$, we estimate the following Tobit model:

\begin{align*}
Y^*_{it} &= \alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t + \epsilon_{it} \\
Y_{it} &= \begin{cases}
          0~\text{if}~ Y_{it}^* \leq 0 \\
          Y_{it}^* ~\text{if}~ 0 < Y_{it}^* < c\\
          c ~\text{if}~ Y_{it}^* \geq c \\
          \end{cases}
\end{align*}

where:

-   $Y_{it}$ is the observed outcome for country $i$ at time $t$, which may be the IMF loan-to-GDP ratio, IMF participation rate, or the number of IMF conditions.
-   $Y^*_{it}$ is the latent continuous variable underlying the observed outcome.
-   $c$ is the upper censoring point, set to 1 for participation rate and $\infty$ otherwise.
-   $\text{USA}_{it}$ and $\text{EUP}_{it}$ represent the propensities for country $i$ to benefit from US and European influence at time $t$, respectively, extracted via principal component analysis.
-   $\mathbf{P}_{it} = \left( \ln(\text{GDP}_{it}), \ln(\text{GDP}_{it})^2, \text{GDPpc}_{it}, \text{GDPpc}_{it}^2 \right)^\top$ includes logged GDP and per capita GDP, along with their squared terms.
-   $\mathbf{Z}_{it}$ is a vector of additional covariates, including per capita GDP growth, reserves as a share of imports, and OECD membership.
-   $\lambda_t$ denotes five-year period fixed effects.
-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^10].

[^10]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$). In such cases, including country fixed effects would lead to inconsistent estimates of the disturbance variance $\sigma$ [@greene2004].

For the IMF loan approval variable, which is binary ($Y_{it} \in \{0, 1\}$), we estimate a Probit model. This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model. The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^11]:

[^11]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.

$$
Y_{it} = 
\begin{cases}
1 & \text{if } Y^*_{it} > 0 \\
0 & \text{if } Y^*_{it} \leq 0
\end{cases}
$$

## Results

We now present the results from estimating the Tobit and Probit models described above, focusing on the effects of US and European influence on IMF lending outcomes. As shown in @tbl-results, stronger ties with Europe are significantly associated with a higher IMF loan-to-GDP ratio, a greater participation rate, and an increased likelihood of loan approval. In contrast, similar ties with the United States do not yield statistically significant effects on these outcomes. The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^12].

[^12]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$. $F$-tests are used for the linear models; for the Probit model, the test corresponds to an asymptotic Chi-squared statistic, which is appropriate given the large sample size ($N = 522$). The $p$-values for the first two models are 0.021 and 0.014, respectively. For loan approval, the difference is nearly significant at the 10% level ($p = 0.122$).

The only favourable treatment associated with American influence, perhaps surprisingly, is a reduction in the number of IMF conditions. In contrast, ties with Europe are not associated with such benefits. This finding echoes the argument by @lutz2014, who suggest that European actors have historically adhered more strictly to economic orthodoxy—most visibly during the Eurozone crisis—despite such policies traditionally being associated with the United States under the so-called “Washington Consensus.” Although the European adoption of neoliberal economic doctrines became especially prominent after the crisis, it is notable that similar patterns are already evident in the pre-2010 data examined here.

```{r}
#| label: tbl-results
#| tbl-cap: Comparative Effects of US and European Influence on IMF Lending Outcomes (1980–2010)
#| tbl-pos: H
load('save/regTable.RData')
resultsTable |> tinytable::theme_tt("resize", width = 1)
```

## Robustness: Continental Preference?

One plausible alternative explanation for the strong association between European influence and favourable IMF lending outcomes is that European powers might be disproportionately favouring fellow European recipients, in contrast to the United States, for which such continental preference may not apply. This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies.

To evaluate this possibility, we examine whether the effect of European influence varies systematically between European and non-European countries. We interact the European influence variable (`EUP Influence`) with a binary indicator for whether the recipient is a European country (`European recipient`), and re-estimate the main models accordingly. The results are presented in @tbl-results-europe.

```{r}
#| label: tbl-results-europe
#| tbl-cap: Effects of US and European Influence on IMF Lending Outcomes (1980–2010), Separating European and Non-European Recipients
#| tbl-pos: H
load('save/regTable_europe.RData')
resultsTable |> tinytable::theme_tt("resize", width = 1)
```

Contrary to the hypothesis of regional favouritism, the results suggest that European recipients do not receive preferential treatment from the IMF or benefit more from their ties to Western European powers. If anything, the estimates indicate that they may even be treated less favourably in some respects. The interaction terms (`EUP Influence` $\times$ `European recipient`), which capture whether European influence has a differential effect for European borrowers compared to non-European ones, are either statistically insignificant or, when significant, indicate a *weaker* effect for European recipients. For instance, in the loan-to-GDP ratio, participation rate, and IMF conditions models, the negative and significant interaction terms suggest that ties with Western European powers are associated with *less favourable* outcomes for European recipients relative to non-European countries with comparable ties.

Furthermore, the `European recipient` variable itself is associated with *less favourable* outcomes, even after controlling for ties to the US and European powers, as well as other covariates. This negative association is statistically significant at the 5% level for both the loan-to-GDP ratio and the number of IMF conditions. These results suggest that, on average, European countries tend to receive *less favourable* treatment from the IMF than non-European countries, despite their typically stronger connections to Western European powers.

These findings suggest that the observed European influence is not driven by a preference for geographically proximate countries, but instead reflects a broader institutional influence over the IMF. In other words, European powers appear to exercise their influence globally, rather than selectively favouring their regional neighbours.

# Case Study

# Conclusion

\newpage

# References {.unnumbered}

::: {#refs}
:::

\appendix

\newpage

# Correlation Matrices {#sec-corr .appendix}

```{r}
#| label: fig-corr-eu
#| fig-cap: Correlation Matrix of European Influence Variables
#| fig-pos: H
#| fig-height: 5
load('save/corr.RData')
corrplot::corrplot(corr_EU, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
```

```{r}
#| label: fig-corr-us
#| fig-cap: Correlation Matrix of US Influence Variables
#| fig-pos: H
#| fig-height: 5
corrplot::corrplot(corr_US, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
```

# Scree Plots {#sec-scree .appendix}

```{r}
#| label: fig-scree-eu
#| fig-cap: Scree Plot of European Influence PCA
#| fig-pos: H
load('save/scree.Rdata')
scree_EU
```

```{r}
#| label: fig-scree-us
#| fig-cap: Scree Plot of US Influence PCA
#| fig-pos: H
scree_US
```