### 2025-08-20 manuscript/manuscript.qmd
Add footnote explaining youssef2023

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
date-format: |
    | DD MMMM YYYY
authors:
  - name: Dianyi Yang
    email: dianyi.yang@politics.ox.ac.uk
    orcid: 0009-0004-4652-3429
    corresponding: true
    affiliations:
      - ref: ox    
  - name: Tong Hu
    orcid: 0009-0003-5234-8164
    affiliations:
      - ref: bw
    equal-contributor: true
  - name: Zhenyi Chen
    orcid: 0009-0008-2910-5187
    affiliations:
      - ref: rmu
    equal-contributor: true
affiliations:
  - id: ox
    name: Department of Politics and International Relations, University of Oxford
  - id: bw
    name: School of International Relations and Diplomacy, Beijing Foreign Studies University
  - id: rmu
    name: School of International Studies, Renmin University of China
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
abstract: "This paper investigates whether the United States or Europe holds greater influence over the International Monetary Fund (IMF)—a longstanding yet underexamined question in the study of global economic governance. While the Fund is widely regarded as dominated by “Western” powers, most existing research either focuses exclusively on American influence or treats the West as a unified bloc, masking important variation between its two key actors. Drawing on panel data from 1980 to 2010, we construct separate measures of US and European influence using principal component analysis of commonly used alignment indicators—UN voting patterns, bilateral trade, and banking exposure—and estimate a series of Tobit and Probit models to assess their impact on four key IMF lending outcomes: loan-to-GDP ratios, participation rates, loan approvals, and conditionality. We find that European influence is strongly associated with higher loan volumes, more frequent participation, and greater likelihood of loan approval, while US influence is not. However, American ties are significantly correlated with fewer conditions attached to loans. Additional analysis shows that European influence is not driven by regional or colonial favouritism: neither European nor African borrowers benefit more from European ties—and may, in fact, fare worse—than other countries with comparable alignment. These findings complicate existing assumptions about IMF decision-making, highlighting the need to distinguish between types of Western influence. They also suggest that Europe wields broader institutional influence over access to IMF resources, while US power manifests more selectively through conditionality. In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
keywords: IMF, US hegemony, Europe
execute:
  echo: false
  warning: false
  error: false
  message: false
---

\newpage

# Introduction

The International Monetary Fund (IMF) and the World Bank[^1] were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004]. While the World Bank specialised in long-term project financing, the IMF was initially established for providing short-term balance-of-payment support and ensuring the stability of the international monetary system [@fund2020; @jensen2004].

[^1]: Familiar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.

Despite IMF's initial technical and narrow mandate, the IMF gradually expanded its role into structural adjustment and domestic economic governance, with loan conditionality often targeting fiscal policy, labour markets, and institutional reforms [@barnett2004]. This evolution pushed the Fund into deeply political and distributional arenas far beyond its original design.

However, despite the far-reaching nature of IMF loan conditionality—often inspired by the tenets of neoclassical economics [@best2007; @clift2021; @chwieroth2015; @nelson2014], such as fiscal austerity, market liberalisation, and privatisation—there is little robust evidence that these prescriptions consistently lead to improved economic outcomes. In many cases, IMF programmes have been associated with worsening economic inequality [@garuda2000; @lang2020; @oberdabernig2013], stagnating growth [@dreher2006; @przeworski2000; @oberdabernig2013], or social unrest [@hartzell2010; @casper2017], raising doubts about the effectiveness and legitimacy of its policy advice.

This performance problem becomes a *legitimacy* issue when compounded by the Fund’s institutional design, which structurally privileges the interests of its most powerful member states [@stone2011]. With a weighted voting system [@fund2020] and governance norms that concentrate decision-making among a handful of advanced economies—particularly those in the West—the IMF has long been criticised as a platform for great power projection [@barnett2004]. Scholars have drawn parallels between the Fund’s influence and forms of informal empire or *neo-colonialism*, where powerful donor states impose conditions on economically weaker recipients through seemingly technocratic institutions [@cain2022; @chen2021; @hickel2020; @moyo2024].

While critiques of Western dominance in the IMF are longstanding, they often treat “the West” as a monolithic bloc. In practice, however, the institutional structure of the IMF reflects a careful transatlantic balancing act. One of the most symbolic manifestations of this arrangement is the long-standing “gentlemen’s agreement” under which the IMF’s Managing Director is always a European, while the World Bank is headed by an American [@keating2024; @weiss2023]. This tacit division of leadership—despite being absent from the formal Articles of Agreement [@fund2020; @WorldBank2021]—raises an underexplored but important question: *who actually wields more influence within the IMF, Europe or the United States?* Addressing this question helps clarify the nature of global economic governance and sheds light on the dynamics within what is often treated as a unified Western core.

The question is straightforward but surprisingly difficult to answer. Whereas the US dominance in the World Bank is undisputed [@clark2020; @wade2002; @woods2003], the IMF presents a more ambiguous case. As the [Literature Review](#sec-litreview) section shows, existing research offers abundant evidence of both American and European influence, yet rarely attempts a direct comparison between the two.

Qualitative approaches to this issue are often hampered by the Fund’s closed-door, consensus-based decision-making [@stone2011, pp.55-56]—precisely the kind of opaque process exemplified by the “gentlemen’s agreement.” On the one hand, the prevalence of informal conventions limits the utility of formal institutional analyses (e.g., voting power or veto rights). On the other, the secretive nature of negotiations means scholars often observe the outcomes, but not the underlying power struggles. While case studies such as Mexico’s bailout or the European debt crisis [@lipscy2018; @copelovitch2010a; @hodson2014] provide compelling narratives, they lack transatlantic comparability and generalisability.

This paper contributes to the literature in three key ways. First, it offers a direct empirical comparison of US and European influence over IMF lending outcomes—an angle surprisingly overlooked in existing work. Second, by disaggregating “the West” into its US and European components, it uncovers variation in how influence is exercised and to what ends. Third, it challenges the assumption of geographic bias, showing that European powers do not disproportionately favour European borrowers. These findings shed new light on the internal politics of the IMF and speak to wider debates on institutional legitimacy, great power contestation, and the geopolitics of global finance.

The structure of this paper is as follows: @sec-litreview provides a review over the literature on the influence of the US and Europe over the IMF.

# Literature Review {#sec-litreview}

International Relations theory, particularly realism, has long emphasised that international organisations reflect underlying distributions of power rather than operating as neutral or autonomous actors. As Mearsheimer famously argued, international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994]. The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a; @stone2011; @vreeland2009]. Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.

In the case of the IMF, there is broad consensus that the institution is dominated by “Western” powers. This is reflected in both historical accounts of its founding and institutional design—especially its weighted voting system, which concentrates power in the hands of the United States and European countries [@kahler2001; @barnett2004; @copelovitch2010]. Given the ideological, geopolitical, and economic proximity between the US and major European states, many studies treat “the West” as a unified bloc and do not disaggregate their respective influences [e.g. @copelovitch2010a; @lipscy2018; @barro2005].

For the strand of literature that examines American and European influence separately, nearly all studies identify the United States as the dominant actor within the IMF [@stone2011, p.51]. Qualitative accounts emphasise the United States’ foundational role in shaping the IMF’s governance structure [@stone2011], the Fund’s practice of making direct enquiries with Washington for major lending decisions [@southard1979; @woods2003], and its unique status as the only country-level veto player, stemming from its position as the largest financial contributor [@woods2003; @stone2011]. The US also benefits from significant informational advantages over other members [@stone2011, p.57] and proximity-based access, with the IMF headquartered in Washington, D.C. [@stone2011, p.57]. Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018]. This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].

Compared to the extensive literature on US dominance in the IMF, research on European influence remains more limited and fragmented. Nevertheless, a growing body of scholarship highlights that European countries do possess considerable institutional and structural advantages that enable them to shape lending outcomes. First, EU countries collectively hold over 30% of IMF voting shares—greater than the US—and, if coordinated, could theoretically counterbalance American veto power [@brandner2009a; @leech2005a; @smaghi2004]. Second, Europe has enjoyed long-standing leadership within the Fund: all IMF Managing Directors since the institution's founding have been European nationals [@kahler2001; @keating2024]. Third, coordination mechanisms such as EURIMF and SCIMF facilitate policy alignment among EU members and amplify their collective voice [@lutz2014]. During the Eurozone crisis, for instance, France and Germany lobbied for IMF involvement, embedding austerity-oriented conditionality that aligned with creditor interests and protected domestic banking sectors [@hodson2014; @kickert2019a; @veron2016]. Scholars have also documented European influence beyond the continent—for example, in the use of IMF concessional lending to sustain post-colonial relationships in Africa, particularly by the UK and France [@moyo2024; @stone2004]. These studies suggest that European actors, when effectively coordinated, are capable of steering lending priorities.

At the same time, existing literature raises questions about the consistency and effectiveness of European influence. Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999]. Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015; @copelovitch2010a]. Moreover, some scholars argue that European influence within the IMF is frequently constrained by the need to align with other G5 powers (refers to the US, Japan, the UK, France, Germany) — particularly the United States — which often necessitates policy compromises and limits Europe’s ability to assert unified preferences [@copelovitch2010]. While some proposals have called for the consolidation of EU seats to strengthen European influence, critics argue this may entrench the dominance of larger EU states without necessarily enhancing representativeness [@leech2005a; @copelovitch2010a]. These institutional tensions may place constraints on Europe’s ability to act decisively, though the extent of such limitations remains contested.

Although many studies document specific cases of European involvement—especially during regional crises—they tend to be case-based, focus on intra-European dynamics, or treat Europe as a secondary actor. In addition, several key works either overlook European influence altogether or conflate it with broader “Western” alignment [@lipscy2018; @barro2005]. As a result, generalisable, cross-national assessments of Europe’s role in shaping IMF outcomes remain scarce. This paper addresses that gap by analytically disaggregating US and European influence within a unified empirical framework. In doing so, it contributes to a more nuanced understanding of transatlantic dynamics in global financial governance—highlighting that Western influence may not be as unified or symmetrical as often assumed.

In principle, quantitative approaches are well suited for directly comparing the influence of different great powers within international organisations. In the IMF context, however, most such studies focus on American influence, with little systematic attention paid to European power. These studies typically proxy the strategic interests of major powers using indicators such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004]. Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009]. Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023][^2]. These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009; @youssef2023]. Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.

[^2]: @youssef2023 also include China as a major shareholder when calculating political influence.

This paper addresses this gap by conducting a disaggregated, quantitative comparison of US and European influence on IMF lending outcomes. By distinguishing between the two major transatlantic actors, we provide new insights into the internal dynamics of power within the IMF and contribute to broader debates over institutional legitimacy, hegemony, and the geopolitics of economic governance.

# Empirical Strategy

## Hypotheses

Since the [literature review](#sec-litreview) has outlined arguments for both US and European influence over IMF decision-making, this section turns to a quantitative comparison of their respective impacts. Accordingly, we formulate our main hypothesis:

\begin{restatable}[$H_{1}$]{hyp}{hypmain}
\label{hyp-main}
The United States and Europe exert different levels of influence on IMF lending decisions.
\end{restatable}

Moreover, prior literature has either combined the United States and Europe under the umbrella of “Western” influence on IMF lending decisions [@lipscy2018], or has examined the influence of one without accounting for the other [@oatley2004; @thacker1999]. To address this gap, we reassess the influence of the United States and Europe separately, while controlling for the presence of the other. Accordingly, we formulate the following two hypotheses:

```{=latex}
\setcounter{hyp}{2}
\begin{restatable}[$H_{2a}$]{subhyp}{hypus}
\label{hyp-us}
The United States exerts a statistically significant influence on IMF lending decisions, controlling for European influence.
\end{restatable}
\vspace{-1em} % remove vertical space
\begin{restatable}[$H_{2b}$]{subhyp}{hypeu}
\label{hyp-eu}
Europe exerts a statistically significant influence on IMF lending decisions, controlling for United States influence.
\end{restatable}
```

## Data

```{r}
#| label: fig-data
dta <- readr::read_rds('data/panel_data.rds')
```

This paper uses the dataset compiled by @lipscy2018, which extends the @barro2005 dataset through 2010. It provides panel data for 130 countries in five-year intervals from 1980 to 2010[^3], including information on IMF borrowing and participation outcomes, economic indicators, and measures of each country's relationship with the United States and major European powers.

[^3]: e.g. 1980-1985, 1985-1990.

## Dependent Variables

In line with @lipscy2018 and @barro2005, we use the following four measures of IMF lending outcomes as our dependent variables:

-   IMF loan to GDP ratio
-   IMF participation rate
-   IMF loan to GDP approval
-   Number of IMF conditions

These variables follow the definitions used in @lipscy2018. The *loan-to-GDP ratio* captures the average size of IMF lending relative to a country's GDP over each five-year interval. The *participation rate* measures the share of months within each interval during which a country was subject to an IMF programme. The *approval indicator* is a binary variable coded 1 if a new IMF programme was initiated during the interval, and 0 otherwise [see also @barro2005]. Finally, the *number of conditions* reflects the total number of performance criteria, benchmarks, and prior actions required for disbursement under IMF programmes [see also @copelovitch2010; @dreher2015].

## Independent Variables

To assess the influence of the United States and European powers on IMF lending decisions, we construct two key independent variables that capture each actor's *propensity* to exert influence in favour of a country's borrowing from the IMF at a given time. These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018]. This approach assumes that both the US and European powers[^4] are more likely to advocate for lending decisions that benefit their close partners.

[^4]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values. This approach reflects the assumption that smaller European countries are unlikely to exert meaningful influence over IMF decisions on behalf of their partners.

Specifically, we construct our independent variables using three indicators commonly interpreted as channels of political and economic affinity, similar to @lipscy2018:

-   UN General Assembly voting similarity[^5]
-   Bilateral trade intensity[^6]
-   Banking sector exposure[^7]

[^5]: Originallly used in @thacker1999.

[^6]: From @barro2005.

[^7]: See @broz2006; @lipscy2017; @oatley2004.

```{r percentage_variance}
load('save/PCA.Rdata')
US_PC1_percentage <- US$sdev[1]^2 / sum(US$sdev^2) * 100
EU_PC1_percentage <- EU$sdev[1]^2 / sum(EU$sdev^2) * 100
```

However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^8]. To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008]. Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^9].

[^8]: See Appendix [-@sec-corr].

[^9]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States). See Appendix [-@sec-scree] for scree plots and additional details.

Our approach differs from @lipscy2018 in that we do not pool the US and European variables together. Instead, we construct two separate measures: US influence and European influence. The PCA is conducted separately for each actor, using the same three indicators described above. The resulting first principal components are then normalised to have a mean of zero and a standard deviation of one, allowing for easier interpretation and comparability of coefficients in subsequent analyses.

## Covariates

The existing literature has identified several factors beyond geopolitical alignment that influence IMF lending decisions or a country's demand for IMF assistance. We include these variables as covariates to account for potential confounding effects. Notably, countries with larger IMF quotas (and thus greater voting power), as well as those with nationals among IMF staff, are expected to receive more favourable treatment from the Fund [@barro2005].

We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997]. To capture these dynamics, we include GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance. International reserves as a share of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support. Finally, a binary variable for OECD membership serves as a proxy for broader institutional and economic differences among countries.

The selection of covariates, including their polynomial transformations[^10], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.

[^10]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.

## Model Specification {#sec-model}

One notable feature of IMF loan outcomes is that they are *left-censored* at zero—that is, loan amounts, participation rates, and the number of conditions are often zero, but cannot take negative values. In such cases, ordinary least squares (OLS) estimation yields biased and inconsistent results. The Tobit model addresses this issue by explicitly modelling the latent continuous variable ($Y_{it}^*$) underlying the censored observed outcomes ($Y_{it}$) [@tobin1958], and is particularly well-suited to our setting [@amemiya1984].

For each non-binary dependent variable $Y_{it}$, we estimate the following Tobit model:

\begin{align*}
Y^*_{it} &= \alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t + \epsilon_{it} \\
Y_{it} &= \begin{cases}
          0~\text{if}~ Y_{it}^* \leq 0 \\
          Y_{it}^* ~\text{if}~ 0 < Y_{it}^* < c\\
          c ~\text{if}~ Y_{it}^* \geq c \\
          \end{cases}
\end{align*}

where:

-   $Y_{it}$ is the observed outcome for country $i$ at time $t$, which may be the IMF loan-to-GDP ratio, IMF participation rate, or the number of IMF conditions.
-   $Y^*_{it}$ is the latent continuous variable underlying the observed outcome.
-   $c$ is the upper censoring point, set to 1 for participation rate and $\infty$ otherwise.
-   $\text{USA}_{it}$ and $\text{EUP}_{it}$ represent the propensities for country $i$ to benefit from US and European influence at time $t$, respectively, extracted via principal component analysis.
-   $\mathbf{P}_{it} = \left( \ln(\text{GDP}_{it}), \ln(\text{GDP}_{it})^2, \text{GDPpc}_{it}, \text{GDPpc}_{it}^2 \right)^\top$ includes logged GDP and per capita GDP, along with their squared terms.
-   $\mathbf{Z}_{it}$ is a vector of additional covariates, including per capita GDP growth, reserves as a share of imports, and OECD membership.
-   $\lambda_t$ denotes five-year period fixed effects.
-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^11].

[^11]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$). In such cases, including country fixed effects would lead to inconsistent estimates of the disturbance variance $\sigma$ [@greene2004].

For the IMF loan approval variable, which is binary ($Y_{it} \in \{0, 1\}$), we estimate a Probit model. This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model. The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^12]:

[^12]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.

$$
Y_{it} = 
\begin{cases}
1 & \text{if } Y^*_{it} > 0 \\
0 & \text{if } Y^*_{it} \leq 0
\end{cases}
$$

# Results

We now present the results from estimating the Tobit and Probit models described above, focusing on the effects of US and European influence on IMF lending outcomes. As shown in @tbl-results, stronger ties with Europe are significantly associated with a higher IMF loan-to-GDP ratio, a greater participation rate, and an increased likelihood of loan approval. In contrast, similar ties with the United States do not yield statistically significant effects on these outcomes. The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^13].

[^13]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$. $F$-tests are used for the linear models; for the Probit model, the test corresponds to an asymptotic Chi-squared statistic, which is appropriate given the large sample size ($N = 522$). The $p$-values for the first two models are 0.021 and 0.014, respectively. For loan approval, the difference is nearly significant at the 10% level ($p = 0.122$).

The only favourable treatment associated with American influence, perhaps surprisingly, is a reduction in the number of IMF conditions. In contrast, ties with Europe are not associated with such benefits. This finding echoes the argument by @lutz2014, who suggest that European actors have historically adhered more strictly to economic orthodoxy—most visibly during the Eurozone crisis—despite such policies traditionally being associated with the United States under the so-called “Washington Consensus.” Although the European adoption of neoliberal economic doctrines became especially prominent after the crisis, it is notable that similar patterns are already evident in the pre-2010 data examined here.

```{r}
#| label: tbl-results
#| tbl-cap: Comparative Effects of US and European Influence on IMF Lending Outcomes (1980–2010)
#| tbl-pos: H
load('save/regTable.RData')
resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
```

# Robustness Checks

While our baseline results indicate that European powers exert substantial influence over IMF lending outcomes, two plausible explanations could account for these patterns without implying broad, institutionalised influence.

First, European powers may favour geographically proximate borrowers — a *continental preference* — resulting in disproportionately favourable treatment of European recipients. Such behaviour might be expected given the EU’s special Troika arrangements with the IMF during the Eurozone crisis [@hodson2014; @pisani-ferry2013]. Second, given the historical legacy of colonialism, particularly in Africa, European powers might be more inclined to support former colonies (colonial preference), especially those with enduring political, economic, or cultural ties [@moyo2024; @stone2004; @stone2011, p.51].

In this section, we test both explanations by interacting our measure of European influence with indicators for European and African recipients, respectively. If either hypothesis holds, we would expect the interaction terms to be positive and significant for the first three lending outcomes — loan-to-GDP ratio, participation rate, and loan approval, indicating stronger European influence in those contexts. As shown below, neither hypothesis is supported: in both cases, the interaction terms are insignificant or negative, and our core finding — that European influence is associated with more favourable IMF lending outcomes — remains robust across the recipient groups. This suggests that European influence is not geographically or historically confined, but instead reflects a global institutional role within the IMF.

## Continental Preference?

One plausible alternative explanation for the strong association between European influence and favourable IMF lending outcomes is that European powers might be disproportionately favouring fellow European recipients, in contrast to the United States, for which such continental preference may not apply. This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies [@hodson2014; @pisani-ferry2013]. Although the Eurozone crisis itself overlaps only slightly with the end of our sample period (1980-2010), and the Troika’s coordinated lending response occurred largely afterwards, we nevertheless test this hypothesis to assess whether similar continental preference dynamics could explain our findings.

To evaluate this possibility, we examine whether the effect of European influence varies systematically between European and non-European countries. We interact the European influence variable (`EUP Influence`) with a binary indicator for whether the recipient is a European country (`European recipient`), and re-estimate the main models accordingly. The results are presented in @tbl-results-europe.

```{r}
#| label: tbl-results-europe
#| tbl-cap: Effects of US and European Influence on IMF Lending Outcomes (1980–2010), Separating European and Non-European Recipients
#| tbl-pos: H
load('save/regTable_europe.RData')
resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
```

Contrary to the hypothesis of regional favouritism, the results suggest that European recipients do not receive preferential treatment from the IMF or benefit more from their ties to Western European powers. If anything, the estimates indicate that they may even be treated less favourably in some respects. The interaction terms (`EUP Influence` $\times$ `European recipient`), which capture whether European influence has a differential effect for European borrowers compared to non-European ones, are either statistically insignificant or, when significant, indicate a *weaker* effect for European recipients. For instance, in the loan-to-GDP ratio, participation rate, and IMF conditions models, the negative and significant interaction terms suggest that ties with Western European powers are associated with *less favourable* outcomes for European recipients relative to non-European countries with comparable ties.

Furthermore, the `European recipient` variable itself is associated with *less favourable* outcomes, even after controlling for ties to the US and European powers, as well as other covariates. This negative association is statistically significant at the 5% level for both the loan-to-GDP ratio and the number of IMF conditions. These results suggest that, on average, European countries tend to receive *less favourable* treatment from the IMF than non-European countries, despite their typically stronger connections to Western European powers.

These findings suggest that the observed European influence is not driven by a preference for geographically proximate countries, but instead reflects a broader institutional influence over the IMF. In other words, European powers appear to exercise their influence globally, rather than selectively favouring their regional neighbours.

## Colonial Preference?

A second possible explanation is that European influence is concentrated in lending to former colonies, particularly in Africa, where historical, linguistic, and economic ties might motivate preferential treatment, as suggested by @stone2004. To assess this, we interact `EUP Influence` with a binary variable for African recipients and re-estimate our main models. The results are reported in @tbl-results-africa.

```{r}
#| label: tbl-results-africa
#| tbl-cap: Effects of US and European Influence on IMF Lending Outcomes (1980–2010), Separating African and Non-African Recipients
#| tbl-pos: H
load('save/regTable_africa.RData')
resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
```

Contrary to the colonial preference hypothesis, the interaction terms (`EUP Influence` $\times$ `African recipient`) are statistically insignificant in most specifications, and when significant (as in the participation rate model), they are *negative* — indicating that ties with Western European powers, if anything, bring fewer benefits to African recipients than to non-African borrowers. Similarly, the `African recipient` variable itself is not consistently associated with more favourable IMF lending outcomes; in fact, in some models (participation rate and conditionality), it is significantly correlated with *less* favourable terms. Taken together, these results suggest that African recipients do not benefit disproportionately from their historical or post-colonial relationships with European powers—and may, on average, receive less favourable treatment than non-African countries with comparable political ties and economic profiles.

These findings parallel our results for European recipients: they indicate that the observed association between European influence and IMF lending is not driven by a concentration of support for countries with historic colonial ties. Instead, European influence appears to be exercised in a globally consistent manner, extending beyond specific regional or historical relationships.

# Case Study

# Conclusion

\newpage

# References {.unnumbered}

::: {#refs}
:::

\appendix

\newpage

# Correlation Matrices {#sec-corr .appendix}

```{r}
#| label: fig-corr-eu
#| fig-cap: Correlation Matrix of European Influence Variables
#| fig-pos: H
#| fig-height: 5
load('save/corr.RData')
corrplot::corrplot(corr_EU, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
```

```{r}
#| label: fig-corr-us
#| fig-cap: Correlation Matrix of US Influence Variables
#| fig-pos: H
#| fig-height: 5
corrplot::corrplot(corr_US, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
```

# Scree Plots {#sec-scree .appendix}

```{r}
#| label: fig-scree-eu
#| fig-cap: Scree Plot of European Influence PCA
#| fig-pos: H
load('save/scree.Rdata')
scree_EU
```

```{r}
#| label: fig-scree-us
#| fig-cap: Scree Plot of US Influence PCA
#| fig-pos: H
scree_US
```